Understanding brain complexity in naturalistic processing in humans

### 2023-02-13 paper/main.tex

\title{High-order dynamic neural correlations reflect naturalistic processing
in humans} \author{Lucy
  L. W. Owen$^1$, and\
  Jeremy R. Manning\textsuperscript{$1, \dagger$}\\
  [0.1in]$^1$Department of Psychological and Brain
  Sciences,\\Dartmouth
  College, Hanover, NH\\
  \textsuperscript{$\dagger$}Address correspondence to
  jeremy.r.manning@dartmouth.edu}

@@

- As complexity of the stimulus increases, decoding accuracy
increases (Fig.~\ref{fig:decode_interpret},  a.).

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/direction_of_field.pdf}
  \caption{\textbf{Direction of the field (adapted from~\citep{Turk13}).} The evolution of fMRI analyses started with
    univariate activation, which refers to
    the average amplitude of BOLD activity evoked by events of an
    experimental condition. Next, multivariate classifiers are trained
    on patterns of activation across voxels to decode distributed
    representations for specific events. The next, resting
    connectivity, is the temporal correlation of one or more seed
    regions with the remainder of the brain during rest. Additionally,
    task-based connectivity examines how these correlations differ by
    cognitive state. Following this increasing trajectory of
    increasing complexity, full connectivity considers all pairwise
    correlations in the brain, most commonly at rest.  Next, dynamic
    full connectivity considers how full connectivity changes over
    time. Continuing this line of reasoning, we expect higher-order network dynamics might provide even richer insights into the neural basis of cognition.}
  \label{fig:direction_of_field}
\end{figure}

>>>>

\title{Understanding brain complexity in naturalistic processing
in humans} \author{Lucy
  L. W. Owen$^1$, and\
  Jeremy R. Manning\textsuperscript{$1, \dagger$}\\
  [0.1in]$^1$Department of Psychological and Brain
  Sciences,\\Dartmouth
  College, Hanover, NH\\
  \textsuperscript{$\dagger$}Address correspondence to
  jeremy.r.manning@dartmouth.edu}

@@

- As complexity of the stimulus increases, decoding accuracy
increases (Fig.~\ref{fig:decode_interpret},  a.).  Replication of
Simony findings.

- As complexity of the stimuli increases, we need fewer components to
decode the same amount (Fig.~\ref{fig:decode_interpret},  b.).

- As complexity of the stimuli increases, more components are required to reach peak decoding accuracy (Fig.~\ref{fig:decode_interpret},  c.).

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/decode_pcs_network.pdf}
  \caption{\textbf{Inflection points by network.} \textbf{a.}
    Inflection point was calculated as explained in
    Fig.~\ref{fig:decode_interpret}, b. Analyses were limited by the
    brain networks (using the Yeoetal11 network parcellation) and
    arranged in increasing order relative to the intact condition.
    \textbf{b. and c.} For the total time in the intact condition, we are plotting the relative inflection points (\textbf{b.}) and corresponding number of components (\textbf{c.}) by network. \textbf{d.} The network parcellation defined by Yeo et al. (2011) is displayed on the inflated brain maps. The colors and network labels serve as a legend for \textbf{a.} and \textbf{d.}}
  \label{fig:decode_pcs_network}
\end{figure}

As complexity of the stimuli increases, decoding accuracy increases
with higher cognitive areas. (Fig.~\ref{fig: decode_pcs_network}).

 If, there is some understanding of the narrative that accumulates over time, we should be able to see that change.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/decode_pcs.pdf}
  \caption{\textbf{Inflection points by thirds.} \textbf{a.}
    Decoding accuracy by number of components not broken into thirds
    (Fig.~\ref{fig:decode_interpret} a.). \textbf{b. and c.}
    Quantifying changes in decoding accuracy across
time. \textbf{b.} Slope of decoding accuracy was calculated by fitting
a regression line for each component/condition for each
third. \textbf{c.}  We also repeated the analysis
(Fig.~\ref{fig:decode_interpret}, b.) to obtain the inflection point
for each condition and for each third. \textbf{d.}  Decoding accuracy by number of components for each third of the scan time. We repeated the same analysis in Fig.~\ref{fig:decode_interpret} a. but breaking the scan time for each condition into 3 intervals.
}
  \label{fig: decode_pcs_thirds}
\end{figure}

- Increases in decoding accuracy with the same number or fewer components for more complex, cognitively rich, conditions.
- Decreases in decoding accuracy for the word-scrambled and rest
condition.

@@

- We trained classifiers using more and more principle components to decode, and compared across condi- tions with varying degrees of cognitive richness.
-We found that as listening conditions become more cognitively rich,
decoding accuracy increased.
-Also, decoding accuracy increased as understanding of the narrative
accumulated over time, in more complex listening conditions.
- Decoding accuracy also increased in higher cognitive areas, in more complex listening conditions.
-We found that as story listening conditions become more complex, more
components are required to decode.
-We also found we could decode better with more impoverished data when
there is the underlying structure of the narrative providing more cognitive richness.
-We posit that as the complexity of our thoughts increases, neural
compression decreases. However, as our thoughts become deeper and richer, more reliable information is available at higher neural compression.
